Predr_Cors

### 2024-03-25 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
pre submission polish

The set of software tools and files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. Training and professional practice guidance for health economists typically provide only superficially coverage of software development topics. However, these capabilities will be critical for achieving proposed methodological innovations at the frontier of the health economics field. Opportunities for field development that closely overlap with CHEM implementation approaches include living health technology assessments [@thokala2023living], machine learning powered health economic analyses [@padula2022machine], open source health-economic models [@Pouwels2022] and utilising Large Language Models (LLMs) to program health economic models [@poirrier2023msr26]. <!-- Potential enhancements to health economic modelling practice include improved model transferability [@munthe2019systematic; @barbieri2010international], greater use of reference models [@Afzali2013], living health technology assessments [@thokala2023living] and leveraging the analytic power of machine learning [@padula2022machine]. -->

There is a significant ethical dimension to software development. For example, recent developments in LLMs mean software developers may confront novel ethical issues including: new security risks; more restrictive open-source licensing; environmentally harmful levels of computing power; and concerns about model explainability and unintended consequences [@ozkaya2023application]. The health economics community should engage more directly with these ethical debates.  The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. Documents such as the World Health Organisation's recent guidance on generative artificial intelligence [@world2024ethics] can be important reference points in helping the health economics discipline navigate relevant ethical topics in software development.

  <!-- Moved -->
 <!-- However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019], despite in principle support from many in the health economics field [@Pouwels2022]. -->
 <!-- Moved -->

>>>>

The set of software tools and files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. Training and professional practice guidance for health economists typically provide only superficial coverage of software development topics. However, these capabilities may be important for achieving a number of proposed methodological innovations for advancing the health economics field. Living health technology assessments [@thokala2023living], machine learning powered health economic analyses [@padula2022machine], open source health-economic models [@Pouwels2022] and the use of Large Language Models (LLMs) to program health economic models [@poirrier2023msr26] all require careful consideration of CHEMs as software development projects. <!-- Potential enhancements to health economic modelling practice include improved model transferability [@munthe2019systematic; @barbieri2010international], greater use of reference models [@Afzali2013], living health technology assessments [@thokala2023living] and leveraging the analytic power of machine learning [@padula2022machine]. -->

There are important ethical dimensions specific to software development that the health economics community should engage with more directly. For example, recent developments in LLMs pose novel ethical issues for software developers including: new security risks; more restrictive open-source licensing; environmentally harmful levels of computing power; and concerns about model explainability and unintended consequences [@ozkaya2023application]. The ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. However, documents such as the World Health Organisation's recent guidance on generative artificial intelligence [@world2024ethics] can be reference points to help identify ethical topics relevant to health economic model software development.
<!-- computational modelling landscape is changing fast [@calder2018computational] and the -->
  <!-- Moved -->
 <!-- However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019], despite in principle support from many in the health economics field [@Pouwels2022]. -->
 <!-- Moved -->

### 2024-03-25 Markdown/Manuscript/Child_RMDs/MS_Tables_1.Rmd
for coauthor feedback draft
pre submission polish

```{r proscons, tab.cap='Implications of choosing commercial modelling or spreadsheet software or an open-source programming language for the computational implementation of health economic models', tab.id = 'proscons', echo = F, results = 'asis'}

@@

                 `Commercial modelling or spreadsheet software` = c("(+) Will include tools to efficienty accomplish many common modelling tasks.",

@@

                                                     "(-) Requirement to manually transfer results to manuscripts through repeated copy and paste actions or transcription from visual inspection increases risk of (undetected) reporting errors.",

@@

                                                        "(ii)	maintaining and transferring (e.g. to new decision problems / contexts) model.",

@@

      dplyr::filter(!`Commercial modelling or spreadsheet software` %in% c("Transparency", "Reusability","Updatability")) %>%

>>>>

```{r proscons, tab.cap='Implications of choosing proprietary software or an open-source programming language for the computational implementation of health economic models', tab.id = 'proscons', echo = F, results = 'asis'}

@@

                 `Proprietary modelling or spreadsheet software` = c("(+) Will include tools to efficiently accomplish many common modelling tasks.",

@@

                                                     "(-) Manual transfer of results to manuscripts through copy and paste or transcription from visual inspection increases risk of (undetected) reporting errors.",

@@

                                                        "(ii)	models are to be maintained and transferred (e.g. to new decision problems / contexts).",

@@

      dplyr::filter(!`Proprietary modelling or spreadsheet software` %in% c("Transparency", "Reusability","Updatability")) %>%

### 2024-03-25 Markdown/Manuscript/Child_RMDs/MS_Tables_3.Rmd
pre submission polish

\renewcommand{\arraystretch}{2}
```{r trucrit, tab.cap='Criteria for assessing the transparency, reusability and updatability of CHEMs', tab.id = 'cpkgs', echo = F, results = 'asis'}
library(kableExtra) 
df <- data.frame(Attribute = c(rep("Transparent",2), rep("Reusable",2), rep("Updtatable",2)),
                 Item = c("T","R","U") %>% purrr::map(~paste0(.x,1:2)) %>% purrr::flatten_chr(),
                 Criterion = c("Software files (executable code, testing procedures and outcomes and non-confidential data) are publicly available.",
                               "It is easy to see who funded, developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use.",
                               "Programming practices facilitate selective reuse of model sub-components.",
                               "Licenses allow anyone to ethically reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works.",
                               "Maintenance infrastructure is in place to support version control and collaboration with model users.",
                               "Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users."))
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      linesep = "\\addlinespace",
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      #kableExtra::pack_rows("Resource requirements implications",1,4, hline_after = T) %>%
      #kableExtra::pack_rows("Ethical implications",5,14, hline_after = T) %>%
      kableExtra::column_spec(1, width = "8em", bold=T) %>% #, latex_valign = "m"
      kableExtra::column_spec(2, width = "3em") %>%
      kableExtra::column_spec(3, width = "28em") %>%
      kableExtra::collapse_rows(columns = 1, latex_hline = "major") 
    # %>%
    #   kableExtra::kable_styling(latex_options = "striped")
      # kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      # kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      # kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)

>>>>

\renewcommand{\arraystretch}{2}
```{r trucrit, tab.cap='Criteria for assessing the transparency, reusability and updatability of CHEMs', tab.id = 'cpkgs', echo = F, results = 'asis'}
library(kableExtra) 
df <- data.frame(Attribute = c(rep("Transparent",2), rep("Reusable",2), rep("Updtatable",2)),
                 Item = c("T","R","U") %>% purrr::map(~paste0(.x,1:2)) %>% purrr::flatten_chr(),
                 Criterion = c("Software files (executable code, testing procedures and outcomes and non-confidential data) are publicly available.",
                               "It is easy to see who funded, developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use.",
                               "Programming practices facilitate independent reuse of model components.",
                               "Licenses allow anyone to ethically reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works.",
                               "Maintenance infrastructure is in place to support version control and collaboration with model users.",
                               "Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users."))
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      linesep = "\\addlinespace",
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      #kableExtra::pack_rows("Resource requirements implications",1,4, hline_after = T) %>%
      #kableExtra::pack_rows("Ethical implications",5,14, hline_after = T) %>%
      kableExtra::column_spec(1, width = "8em", bold=T) %>% #, latex_valign = "m"
      kableExtra::column_spec(2, width = "3em") %>%
      kableExtra::column_spec(3, width = "28em") %>%
      kableExtra::collapse_rows(columns = 1, latex_hline = "major") 
    # %>%
    #   kableExtra::kable_styling(latex_options = "striped")
      # kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      # kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      # kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)
